---
title: "Horsepower and Fuel Efficiency by Cylinder Count"
author: Owen Lin
date: "2026-09-23"
---

Setting up my environment:

```{r}
#| message: false
#| warning: false
library(tidyverse)
```

Using the R's built in mtcars data, let's examine the cylinder count, horsepower, and fuel efficiency of each car.

```{r}
#Pulling relevant information about cars
cars <- mtcars |>
  tibble::rownames_to_column("car") |>
  select(car, cyl, hp, mpg) |>
  arrange(cyl)

knitr::kable(cars)
```

What is the average horsepower and fuel efficiency for each cylinder count?

```{r}
#Finding average horsepower and fuel efficiency for cars of each cylinder count
cyl_avg <- cars |>
  group_by(cyl) |>
  summarise(
    avg_hp = mean(hp),
    avg_mpg = mean(mpg)
  )
```

Presenting our findings:

```{r}
#| label: tbl-avg-horsepower-mpg-by-cylinder
#| fig-cap: "Average horsepower and MPG by cylinder count."
knitr::kable(
  cyl_avg,
  digits = 1,
  col.names = c("Cylinders", "Avg. horsepower", "Avg. MPG")
)
```

As cylinder count increases, average horsepower increases and average fuel efficiency decreases.

# Source

Data: [mtcars](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html),
R datasets package (GPL-2 | GPL-3). Originally from Motor Trend US magazine (1974),
via Henderson & Velleman (1981), *Biometrics* 37: 391–411.